\documentclass[10pt]{article}
\usepackage{times}
%Sets Times, 10pt.

\usepackage[letterpaper]{geometry}
%Sets 8.5x11 - Standard U.S. Letter

\geometry{top=1.0in, bottom=0.5in, left=0.5in, right=0.5in}
%Sets Margins

\usepackage{setspace}
\singlespacing
%Sets single spacing.
\usepackage{parskip}
\parskip=0\baselineskip \advance\parskip by 0pt
%Removes space between paragraphs.

\usepackage{multicol}
%Sets two columns.

\usepackage{amssymb}
%Provides symbols.

\usepackage{fancyhdr}
%This package allows for the formatting of headers and footers
\pagestyle{fancy}
%This allows for all of the pages to be formatted with the headers and footers defined below.
\chead{Duncan Kubalak - Chapter 6 Notes}
\cfoot{}
\renewcommand{\headrulewidth}{0pt}
%Creates a centered heading

\usepackage[english]{babel}
%Sets document language as American English, allows document to format punctuation (quotes, hyphens, etc.) to U.S. English Standards.

\begin{document}
\begin{flushleft}

\begin{multicols}{2}
%Creates two columns, change the number of columns or remove entirely.

\begin{itemize}
\item British planned in 1778 to invade south a second time.
\item[$\vartriangleright$] The Continental army grew by 5,000 in South Carolina to defend Charleston.
\item[$\vartriangleright$] From the Citadel (a Continental army fort) Americans bombarded the British with all they had (makeshift - broken glass \& pickaxes). British responded with ships \& cannons.
\item[$\vartriangleright$] Citadel eventually fell after heavy casualties on both sides, Americans surrendered, British returned to N.Y.
\item Cornwallis (2\textsuperscript{nd} in command to Clinton) took over the fight for the British in the south. His surrender @ Yorktown would end the war in 1781.
\item Francis Marion ``swamp fox'' Leader of Guerrilla fighting in South Carolina. Harassed British during 2\textsuperscript{nd} Southern Campaign. 
\item[$\vartriangleright$] Battle of King's Mountain: Battle between Revolutionaries and Loyalists on the border of the Carolinas. Revolutionary troops defeated Loyalists.
\item Regular continental army, led by Gates was not doing well against the British. Washington replaced Gates with Nathaniel Greene in 1780. Greene eased the strain of starving troops \& partnered with local Native American tribes. 
\item[$\vartriangleright$] Greene used a strategy of attrition: wear the British army out by making them chase his small, nimble army all over the south.
\item[$\vartriangleright$] The courage presented by the outnumbered American troops in battle also scared the British, eventually leading Cornwallis to retreat to Virginia.
\item In 1780 Gen. Benedict Arnold (one of Washington's prot\'eg\'es) defected to the British. Arnold's plot to surrender the fort (@ West Point in N.Y.) to the British failed, but his treason saddened Washington \& lowered morale.
\item Soon after, French aid arrived
\item[$\vartriangleright$] French Gen. Rochambeau wanted to attack Cornwallis in the south, \& Washington wanted to take back N.Y. The French were already in Chesapeake, so Washington reluctantly agreed.
\item At the first sign of an oncoming attack, Cornwallis moved his troops to Yorktown (bad choice).
\item[$\vartriangleright$] French navy was already in place in Chesapeake bay.
\item[$\vartriangleright$] Clinton realized too late that something was happening, and sent a weak rescue party to help Cornwallis.
\item[$\vartriangleright$] French Gen. DeGrasse fought off the rescue party with no problem.
\item[$\vartriangleright$] Washington's troops barraged the redcoats with constant artilery, deafening them, and preventing them from fighting back properly towards the French.
\item Cornwallis soon gave up \& surrendered, effectively ending the war, \& winning it for the colonists.
\item[$\vartriangleright$] Some scattered fighting continued throughout the colonies, until the British parliament declared it was over.
\item Now diplomats had to preserve independence.
\item[$\vartriangleright$] Treaty of Paris, 1783: Ended the rev. war, and secured independence for America.
\item[$\vartriangleright$] Benjamin Franklin, John Adams, and John Jay were sent to Paris.
\item[$\vartriangleright$] Negotiation continued for over a year, a vague treaty gave a lot of land to the Americans, disregarding the Natives who owned it and lived there.
\item America became a Republic, A nation in which supreme power resides in the citizens who elect representatives to govern them.
\item With the country established, the principles set up in the declaration of independence had to be enforced.
\item[$\vartriangleright$] Many supported ``Freedom of Conscience,'' but not all supported the separation of church \& state. Virginia had used tax money to pay for churches; the ``Statute of religious freedom'' put an end to this.
\item A belief in social equality was important. Many colonial-era rules \& social norms were reformed.
\item[$\vartriangleright$] Reformed: Primogeniture - Eldest son inherits entire estate of father.
\item[$\vartriangleright$] Reformed: Entail - Prevents land \& property from being divided, sold, or given away. (This had led to the aristocracy in England).
\item Some women had fought in the war, taking over for fallen men on the battlefield.
\item[$\vartriangleright$] Mary Ludwig - One of many women named ``Molly Pitchers'' who had carried water to cool down cannons in battle. Ludwig was married to a soldier @ Ft. Monmouth.
\item[$\vartriangleright$] Ludwig \& Margaret Corbin did engage in military combat. Corbin later applied for a government pension.
\item America's political \& social 1337 now rethought the role of women in government.
\item[$\vartriangleright$] Out with colonial belief that ``women lack the ability to reason and make moral, rational decisions.'' Thanks, Enlightenment!
\item[$\vartriangleright$] Republican womanhood - A role for mothers that became popularized following the revolution. It stressed women's importance in instructing children in republican virtues, such as patriotism \& honor.
\item Esther DeBerdt Reed - Wealthy Philadelphia matron who helped to raise money for the American army. Women who, like Reed, expressed their patriotism through public action were accused of ``overstepping the boundaries of their gender'' by engaging in ``unfeminine behavior.''
\item[$\vartriangleright$] Reed defended the activism of herself and many other colonial women in ``The Sentiments of an American Woman.'' It compared many prominent women in the American Revolution 
\item The war resulted in educational progress in the colonies; by 1789 Massachusetts required every town to provide free public education to all children (girls and boys).
\item[$\vartriangleright$] Before this women only learned from their parents, \& they probably only learned ``domestic skills,'' rather than mathematics or politics.
\end{itemize}

\end{multicols}

\end{flushleft}
\end{document}
