\documentclass[10pt]{article}
\usepackage{times}
%Sets Times, 10pt.
\usepackage[letterpaper]{geometry}
%Sets 8.5x11 - Standard U.S. Letter
\geometry{top=0.75in, bottom=0.5in, left=0.5in, right=0.5in}
%Sets Margins
\usepackage{setspace}
\singlespacing
%Sets single spacing.
\usepackage{parskip}
\parskip=0\baselineskip \advance\parskip by 0pt
%Removes space between paragraphs.
\usepackage{multicol}
%Sets two columns.
\usepackage{amssymb}
%Provides symbols.
\usepackage{fancyhdr}
%This package allows for the formatting of headers and footers
\pagestyle{fancy}
%This allows for all of the pages to be formatted with the headers and footers defined below.
\chead{Duncan Kubalak - Chapter 3 Notes}
%Creates a centered heading
\cfoot{}
%Ensures that no footer will be created, as sometimes a page number is displayed automatically
\renewcommand{\headrulewidth}{0pt}
%Redefines the predefined header command, and throws an error if it has not been defined.
\usepackage[english]{babel}
%Sets document language as American English, allows document to format punctuation (quotes, hyphens, etc.) to U.S. English Standards.

\begin{document}
\begin{flushleft}
\begin{multicols}{2}
%Creates two columns, change the number of columns or remove entirely.
\begin{itemize}
 \item Mayflower Compact – Social contract created in the name of survival.
 \item[$\vartriangleright$] 102 settlers – 41 signed (all men)
 \item[$\vartriangleright$] Most settlers were critics of English gov’t.
 \item Massachusetts bay colony – puritans, seeking refuge from religious and political persecution.
 \item[$\vartriangleright$] Led by John Winthrop – Lawyer 
 \item[$\vartriangleright$] Better than Jamestown or Plymouth – more similar to England, easier to survive.
 \item[$\vartriangleright$] Colonists were well-prepared, sent some people ahead to make sure the land was ready.
 \item[$\vartriangleright$] Attracted thousands of settlers – “Great Migration” (Ended with Oliver Cromwell \& Puritan army in England).
 \item[$\vartriangleright$] VA colonists – motivated by money; Pilgrims – religious persecution; MA Bay – ``model Christian society.''
\item[$\vartriangleright$] MBC traditional English communities – houses built around a village green, church nearby, surrounded by farmland.
 \item[$\vartriangleright$] In Great Migration people immigrated as families, so the gender ratio was not skewed.
 \item[$\vartriangleright$] Cool temperatures, clean drinking water – healthier place to live than England.
 \item[$\vartriangleright$] Family – “little commonwealth” – hierarchy w/in a family; chain of command. Husband: be respectful to and care for wife. Wife: follow orders, do chores, help other wives in the community.
 \item[$\vartriangleright$] Femme Couverte – French: “Covered Woman” – married woman, limited legal status.
 \item[$\vartriangleright$] Not all colonists were equal. Winthrop though it natural for some to be rich and others poor. “Some high/eminent in power/dignity, others mean/in subjugation.” (Women, children, servants, young men, adult men w/ no property owed obedience to others)
 \item[$\vartriangleright$] Church participation was required for any political activity.
 \item[$\vartriangleright$] In first decades of colony existence little religious tolerance – no reason to welcome others (go elsewhere).
 \item[$\vartriangleright$] Winthrop \& co. aggressive towards Quakers (Protestant pacifists, believed in gender equality).
 \item[$\vartriangleright$] Roger Williams – puritan minister banished from MBC for criticizing religious rules \& gov’t. (Condemned seizure of Native land, and laws requiring church attendance). Later joined Narragansett, his congregation came to find him, they established Providence where all were welcome. – Now Rhode Island. First separation of church and state.
 \item[$\vartriangleright$] Anne Hutchinson – Immigrated with her family, to Mass., she was very educated, didn’t like the strict rules imposed by religious officials. Many who didn’t have rights looked up to her, she taught them. (She also didn’t have rights by their laws.) Arrested and brought to trial. Forced to stand throughout the trial, although she was pregnant. Accidentally stated she had spoken with God, banished as a heretic.
 \item Pequot War – 1636 Natives driven from their land, population destroyed by MB and Conn. colonists.
 \item Metacom – Wampanoag chief who led resistance against colonists, used guerilla tactics, eventually assassinated.
 \item Puritan revolution in England – some immigrants left MBC to fight, few new immigrants came.
 \item MBC lost charter from England (K. Charles) then, northern colonies were combined into “megacolony” Dominion of New England (K. James II) This included Mass, RI, Conn, Plymouth, NJ, NY.
 \item In Salem, several women fell ill, “bewitchment,” many accused of “bewitching” them, tried, jailed, and 19 executed.
 \item Pre-1650 English and Dutch settlers cooperated, this ended when both vied for control of transatlantic trade routes and land in Conn. Valley.
 \item New Netherland became New York without violence, b/c settlers thought life under English would be same as life under Dutch.
 \item Jacob Leisler – German merchant led revolution against English gov’t in NY. Took control, removed Dominion officials, declared them and his enemies as enemies of Protestantism itself. Executed as a traitor.
 \item William Penn – English Quaker who founded Pennsylvania. (Loyal to the king, son of admiral, granted charter for colony)
 \item[$\vartriangleright$] Quakers – egalitarian society – plain dress, didn’t remove hats for “betters,” gender equality.
 \item[$\vartriangleright$] All were welcome in Quaker society, Penn strived for peaceful coexistence with Indians, all land purchased fairly.
 \item[$\vartriangleright$] Philadelphia became a great center of shipping and commerce – at a price to the original values of Pennsylvania.
 \item Carolina colony established where French, Spanish, Dutch, and Natives all inhabited. Founded by 8 colonizers for England.
 \item[$\vartriangleright$] Philosopher John Locke helped establish Carolina constitution, society comprised of Landowners, Yeomen (independent farmers), and Serfs (agricultural laborers bound to work for their landlords.)
 \item[$\vartriangleright$] People were not willing to travel thousands of miles to become serfs.
 \item[$\vartriangleright$] Locke later became famous for his writing about freedom and human rights.
 \item Georgia – 100 years after Jamestown colony, last of the 13 colonies is established in the south.
 \item[$\vartriangleright$] James Oglethorpe – wealthy social reformer, looking for a new moral life for people imprisoned for debts.
 \item[$\vartriangleright$] K. George II wanted to create a buffer between Spanish in Florida and the valuable rice production in South Carolina. (Military service was made mandatory for every man in Georgia – poor of Georgia would protect rich of SC).
 \item[$\vartriangleright$] Oglethorpe ensured hard work by banning alcohol, slavery, and keeping land grants small.
 \item[$\vartriangleright$] Most immigrants were not impoverished, indebted, instead they were middle-class English or South Carolinians looking for more land.
 \end{itemize}
\end{multicols}
\end{flushleft}

\end{document}
