\documentclass[10pt]{article}
\usepackage{times}
%Sets Times, 10pt.

\usepackage[letterpaper]{geometry}
%Sets 8.5x11 - Standard U.S. Letter

\geometry{top=0.75in, bottom=0.5in, left=0.5in, right=0.5in}
%Sets Margins

\usepackage{setspace}
\singlespacing
%Sets single spacing.
\usepackage{parskip}
\parskip=0\baselineskip \advance\parskip by 0pt
%Removes space between paragraphs.

\usepackage{multicol}
%Sets two columns.

\usepackage{amssymb}
%Provides symbols.

\usepackage{fancyhdr}
%This package allows for the formatting of headers and footers
\pagestyle{fancy}
%This allows for all of the pages to be formatted with the headers and footers defined below.
\chead{Duncan Kubalak - Chapter 7 Notes}
%Creates a centered heading
\cfoot{}
%Ensures that no footer will be created, as sometimes a page number is displayed automatically
\renewcommand{\headrulewidth}{0pt}
%Redefines the predefined header command, and throws an error if it has not been defined.

\usepackage[english]{babel}
%Sets document language as American English, allows document to format punctuation (quotes, hyphens, etc.) to U.S. English Standards.

\begin{document}
\begin{flushleft}

\begin{multicols}{2}
%Creates two columns.

\begin{itemize}
\item In May of 1787 George Washington gathered a group of people to draft a constitution in Philadelphia.
\item[$\vartriangleright$] Delegates came from each of the 13 colonies.
\item[$\vartriangleright$] Most of the delegates were nationalists (Americans who preferred strong central government to small, limited government of the \textit{Articles of Confederation}) yet they did not all agree on a strategy.
\item[$\vartriangleright$] Some wanted to revise the \textit{Articles} whereas others wanted to replace them entirely.
\item[$\vartriangleright$] Edmund Randolph introduced the \textit{Virginia Plan}, a list of 14 proposals by the VA delegation for creating a more powerful central government and giving states proportional representation with a bicameral legislature (representatives divided into two separate assemblies).
\item James Madison (VA planter and political theorist ``father of the constitution,'' fourth president of U.S.) was the one who made the VA plan into the government we have today.
\item[$\vartriangleright$] Madison's plan described a gov't with 3 distinct branches (Legislative, Executive, Judicial) to replace the congress of the time, which did all three. This way one group could not wield too much power.
\item[$\vartriangleright$] Madison's plan also included the power of veto: the right of one branch of gov't to reject the decisions of another branch.
\item There was a lot of controversy w/in the delegation over Madison's VA plan.
\item[$\vartriangleright$] Large states supported the plan as it would give them more power in legislation based on their large populations.
\item[$\vartriangleright$] Small states started supporting an alternative, the \textit{New Jersey Plan}, which would give all states an equal vote (w/ a unicameral rather than a bicameral legislature).
\item[$\vartriangleright$] The Great Compromise: satisfied both sides w/ a bicameral legislature but w/ equal rep. in the senate, and proportional rep. in congress.
\item The last big dispute over representation: which Americans count as part of a state's population?
\item[$\vartriangleright$] North wanted slaves to be counted as property, not as people b/c they had more free people \& would get more representation.
\item[$\vartriangleright$] Southern states (many of which were up to $\frac{1}{3}$ slaves) wanted slaves to be counted as people (just w/o rights) so \textit{they} would get more rep. in congress.
\item[$\vartriangleright$] The three-fifths compromise: (defied reason but made political sense) three in five slaves would be counted in a state's population for the house of reps.
\item The only other issues that really mattered to the delegates @ that point was establishing a system of checks and balances.
\item[$\vartriangleright$] The president had executive powers (commander in chief, responsible of foreign affairs, etc.)
\item[$\vartriangleright$] Congress had to offset these powers w/ the power to declare war \& raise an army. They also had the \textit{power of the purse} which would give them the power to tax ppl \& use the money freely.
\item[$\vartriangleright$] One of the most clumsy compromises: the electoral college. (A body of electors chosen by the states to elect the president \& V.P.) States could select a number of electors equal to the number of it's senators \& Congressional reps.
\item[$\vartriangleright$] Delegates thought ppl in the states wouldn't have knowledge of candidates from other states, so they designed a system wherein candidates would be voted upon by electors who were their political colleagues. 
\item Ratifying Conventions: meetings in states to agree on whether they would ratify the constitution in that state.
\item In order to gather support for an act as bold as changing a form of gov't the pro-constitution ppl needed to change the language.
\item[$\vartriangleright$] The Nationalists became \textit{federalists}, a name which although it still suggested a support for centralized government, put more emphasis on state's rights.
\item[$\vartriangleright$] The anti-constitution forces were then cheated out of a name, \& became the ``anti-federalists.'' (They believed a strong central gov't was a danger to American liberties and rights.) The label, however, seemed like they were negative, pessimistic, \& w/o their own agenda.
\item[$\vartriangleright$] Urban, NE areas were more Federalist, while more Rural, frontier areas believed in Anti-federalist principles.
\item Hamilton, Madison, and Jay wrote the \textit{Federalist Papers}, which would convince Americans to accept the new form of gov't.
\item Gov. Morris inspired delegates to think as a country, rather than as residents of specific states, regions, or towns.
\item NY had an intense election, but voted federalist b/c the new ogv't was already \textit{Fait Accompli} (cuoldn't be reversed) out side of the state.
\item After Senators and Congressmen had been elected, it was time for a presidential election.
\item[$\vartriangleright$] G. WA din't expect to run for president, but he was elected because he was one of the only candidates with a national reputation.
\item[$\vartriangleright$] he got the electoral vote right away, \& John Adams (second president) was elected as the VP.
\item[$\vartriangleright$] WA established a cabinet, \& gathered them often to discuss matters of the electoral branch of gov't. (Hamilton = trasury, Knox = war, Randolph = attorney gen. \& Jefferson = Secretary of State.)
\end{itemize}

\end{multicols}

\end{flushleft}
\end{document}
